---
title: "Analysis"
author: "Andrew Redd, PhD."
date: "R Bootcamp 2021"
output:
  ioslides_presentation:
    logo: images/U_Health_stacked_png_red.png
    theme: paper
    css: style.css
    keep_md: no
    md_extensions:
    - +task_lists
    - +multiline_tables
    includes:
      in_header: timer.htm
    standalone: yes
    widescreen: yes
---

```{r setup, include=FALSE}
library(knitr)
library(printr)
library(tidyverse)
library(magrittr)
knitr::opts_chunk$set(echo = TRUE, error=TRUE
                     , class.error='error'
                     , class.warning='warning'
                     , class.message='message'
                     , cache=FALSE)
options( repos="https://cran.rstudio.org"
       , width=60
       )
```

# Data

## Bordeaux Wine Data
```{r}
bordeaux <- readRDS("data/bordeaux.rds")
```

### Variables
* `Wine` - the name of the wine.
* `Price` - The price of the wine in pounds sterling(&pound;)
* `ParkerPoints` - the rating out of 100 given by [Robert Parker](https://www.robertparker.com/).
* `CoatesPoints` - the rating out of 20 given by [Clive Coates](http://www.clive-coates.com/)
* `P95andAbove` - a dummy variable, 1 if `ParkerPoints>=95`
* `FirstGrowth` - inditcator of if the wine is a [first growth](https://en.wikipedia.org/wiki/First_Growth)
* `CultWine` - indicator of if the wine is a [cult wine](https://en.wikipedia.org/wiki/Cult_wine)
* `Pomerol` - indicator for if the wine is from [Pomerol](https://en.wikipedia.org/wiki/Pomerol), France
* `VintageSuperstar` - indicator if the wine is a superstar


# Hypothesis Tests

## T-test

Test if there is a difference in price for first growth wines.
```{r}
(fg.test <- t.test(Price~FirstGrowth, data=bordeaux))
```

-----
:::{.terminology}
Formulas, created with a `~`, represent relationships. They can 

* be one sided `~x`
    - Lambda functions `~log(.+1)`
* or two sided `y ~ x + z`
    - specify relationships or models
* often include functions
    - `y ~ x + log(z)`
    - `y ~ x + poly(z,3)`, a polynomial fit of degree 3 on `z`
* have special syntax
    - interaction `y ~ x:z`
    - crossing `y ~ a*b` is equal to `y ~ a + b + a:b`
    - nesting `y ~ a + b %in% a` or equivalently `y~a/b`
:::

## Getting usable results from a model {.build}

The `fg.test` object is a `r class(fg.test)` object, 
which prints nicely but what if we want to include this in our table 1?

Try these:
```r
str(fg.test) # get the underlying structure of the object.
glimpse(fg.test) # alternative to str that handles some objects better. 
fg.test$p.value
getElement(fg.test, 'p.value')
fg.test[['p.value']]
```


# Linear Models



## Wine Model

We will rely on the normal approximation for proportions.

```{r}
model <- lm( Price ~ . - Wine - P95andAbove, data=bordeaux)
model
```
Not really useful.

## Formula Creation Model

::: {.concept}

Formula Subtraction

: `Price ~ . - Wine - P95andAbove` should be read as

    > "Model Price by all variables **except** Wine and P95andAbove."

:::

## Summarizing Models

```{r}
(model.summary <- summary(model))
```

## Exercise: Try the following

:::::: {.columns-2}
:::{.column}
#### Extracting Parts of the model

```r
coef(model)
coef(summary(model))
deviance(model)
formula(model)
residuals(model)
```
:::
:::{.column}
#### Operations on models

```r
summary(model)
plot(model)
predict(model)
vcov(model)
anova(model)
aov(model)
```
:::
::::::

:::{#modelstimer .timer onclick="startTimer()" warnafter=30}
5:00
:::

## Residual plots

```{r h=4,w=6}
plot(model, 1) #< Residuals vs. Fitted  | Goodness of fit
```

## Residual plots

```{r}
plot(model, 2) #< Normal Q-Q            | Appropriate error model 
```

## Residual plots

```{r}
plot(model, 3) #< Scale-Location        | Homoscedasticity
```

## Residual plots

```{r}
plot(model, 5) #< Residual vs. Leverage | Influential points
```
